Functions to make plain R output look nicer

### 2025-02-09 text/25/pretty-functions.Rmd
Finalized pretty-functions

---
title: "New RMarkdown program"
author: "Steve Simon"
date: "`r Sys.Date()`"
output: html_document
---

@@

```{r}
pretty_mean <- function(d, v) {
  d |>
    summarize(n_missing=sum(is.na({{v}}))) -> d1
  d |>
    filter(!is.na({{v}})) |>
    summarize(
      across({{v}}, 
        list(
          mean=mean, 
          sd=sd,
          min=min,
          max=max))) |>
    bind_cols(d1)
}
```

>>>>

---
title: "Functions to make plain R output look nicer"
author: "Steve Simon"
source: new
date: "2025-02-09"
categories:
- Blog post
tags:
- R programming
output: html_document
page_update: complete
---

@@

[ref10]: https://stat.ethz.ch/R-manual/R-devel/library/base/html/summary.html
[ref11]: https://dplyr.tidyverse.org/reference/group_by.html

```{r}
pretty_mean <- function(d, v) {
  # d is a data frame or tibble, v is a variable in d.
  d |>
    summarize(n_missing=sum(is.na({{v}}))) -> d1
  d |>
    filter(!is.na({{v}})) |>
    summarize(
      across({{v}}, 
        list(
          mean=mean, 
          sd=sd,
          min=min,
          max=max))) |>
    bind_cols(d1)
}
```

Using tidyverse functions [inside a loop or function][ref12] is tricky. With an argument that represents a variable name, you either enclose the argument in double curly braces ("embrace" is the term they use) or you specify the argument inside of .data[[]].

The [across][ref13] function, part of the dplyr/tidyverse libraries, allows you to select specific variable(s) for further operations within dplyr functions like summarize. I have had some difficulty with this function, but it seems to work well here.

The [bind_cols][ref14] function, part of the dplyr library, combines two data frames or tibbles side by side.

[ref12]: https://dplyr.tidyverse.org/articles/programming.html
[ref13]: https://dplyr.tidyverse.org/reference/across.html
[ref14]: https://dplyr.tidyverse.org/reference/bind_cols.html

Here is the code you would write without the pretty_mean function.

```{r}
penguins |>
  summarize(
    body_mass_g_mean=mean(body_mass_g, na.rm=TRUE),
    body_mass_g_sd=sd(body_mass_g, na.rm=TRUE),
    body_mass_g_min=min(body_mass_g, na.rm=TRUE),
    body_mass_g_max=max(body_mass_g, na.rm=TRUE),
    n_missing=sum(is.na(body_mass_g)))
```

Here is the code with the pretty_mean function.

```{r}
penguins |>
  pretty_mean(body_mass_g)
```

Writing functions like these only make sense if you find yourself doing the same sort of thing three or more times in one program. There is a nice side effect of writing the function. If you decide that you want a different appearance or you want to round some of your results or you want to feed the data into a [nicely formatted table][ref15], it only takes a small modification in one spot.

[ref15]: https://cran.r-project.org/web/packages/insight/vignettes/display.html
